---
title: "P8130_Final_Project"
author: "Zilin Huang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(vtable)
library(MASS)
library(corrplot)
library(glmnet)
```

```{r}
data = read.csv("./Project_2_data.csv") |>
  mutate(
    Race = case_match(
      Race,
      "White" ~ 0,
      "Black" ~ 1,
      "Other" ~ 2
  ),
  Marital.Status = case_match(
      Marital.Status,
      "Married" ~ 0,
      "Divorced" ~ 1,
      "Single " ~ 2,
      "Widowed" ~ 3,
      "Separated" ~ 4
  ),
  T.Stage = case_match(
      T.Stage,
      "T1" ~ 0,
      "T2" ~ 1,
      "T3" ~ 2,
      "T4" ~ 3
  ),
  N.Stage = case_match(
      N.Stage,
      "N1" ~ 0,
      "N2" ~ 1,
      "N3" ~ 2
  ),
  X6th.Stage = case_match(
      X6th.Stage,
      "IIA" ~ 0,
      "IIIA" ~ 1,
      "IIIC" ~ 2,
      "IIB" ~ 3,
      "IIIB" ~ 4
  ),
  differentiate = case_match(
      differentiate,
      "Poorly differentiated" ~ 0,
      "Moderately differentiated" ~ 1,
      "Well differentiated" ~ 2,
      "Undifferentiated" ~ 3
  ),
  A.Stage = case_match(
      A.Stage,
      "Regional" ~ 0,
      "Distant" ~ 1,
  ),
  Estrogen.Status = case_match(
      Estrogen.Status,
      "Negative" ~ 0,
      "Positive" ~ 1,
  ),
  Progesterone.Status = case_match(
      Progesterone.Status,
      "Negative" ~ 0,
      "Positive" ~ 1,
  ),
  Grade = case_match(
    Grade,
    "1" ~ 1,
    "2" ~ 2,
    "3" ~ 3,
    " anaplastic; Grade IV" ~ 4
  )
  ) 
```

Include a descriptive table with summary statistics for all variables.

```{r}
st(data)
```

Explore the distribution of the outcome and consider potential transformations (if necessary):

`Tumor.Size` needs log transformation:

```{r}
fit1 = lm(Survival.Months ~ Tumor.Size, data = data)
boxcox(fit1, lambda = seq(-3, 3, by = 0.25))

data =
  data |> 
  mutate(Tumor.Size = log(Tumor.Size))

data |>
  ggplot(aes(x=Tumor.Size)) + geom_histogram()
```

Examine the pairwise relationship between the independent variables 1-14.

```{r}
# pairs(data[, -((ncol(data)-1):ncol(data))])
cor(data[, -((ncol(data)-1):ncol(data))])
corrplot(cor(data[, -((ncol(data)-1):ncol(data))]), type = "upper", diag = FALSE)
```

`Regional.Node.Positive` and `N.Stage`, `Tumor.Size` and `T.Stage`, `Differentiate` and `Grade` are highly linearly correlated, and we choose to manually filter out one variables from each of the three paris for the regression modeling process.

### Model Against Survival Months

```{r}
# The filtering step
data_lin = data |>
  dplyr::select(-Status, -Regional.Node.Examined, -T.Stage, -Grade)
```

First model is based on Forward Selection.

```{r}
mult.fit <- lm(Survival.Months ~ ., data = data_lin)
intercept_only <- lm (Survival.Months ~ 1, data = data_lin)
lin_1 <- step(intercept_only, direction = "forward", scope = formula(mult.fit))
```
```{r}
par(mfrow = c(2,2))
plot(lin_1)
```

Second model is based on LASSO.

```{r}
lambda_seq <- 10^seq(-3, 0, by = .1)
set.seed(2022)
cv_object <- cv.glmnet(as.matrix(data_lin[,1:11]), data_lin$Survival.Months,lambda = lambda_seq, nfolds = 5)

lin_2 <- glmnet(as.matrix(data_lin[,1:11]), data_lin$Survival.Months, lambda = cv_object$lambda.min)
coef(lin_2)
```

Third model is based on variables with demographic information, with their interaction terms added.

```{r}
lin_3 <- lm(Survival.Months ~ Age + Race + Marital.Status + Age*Race + Age*Marital.Status + Race*Marital.Status + Age*Race*Marital.Status, data = data_lin)
summary(lin_3)

anova(lin_3)
```

```{r}
# Improved 
lin_3_0 <- lm(Survival.Months ~ Race + Marital.Status + Race*Marital.Status, data = data_lin)
```

```{r}
par(mfrow = c(2,2))
plot(lin_3_0)
```

Fourth model is based on various Physiological indicators, filtered by ANOVA comparisons.

```{r}
# prefer lin_4_0 without X6th.Stage
lin_4 <- lm(Survival.Months ~ N.Stage + X6th.Stage + A.Stage + Tumor.Size + differentiate, data = data_lin)
summary(lin_4_0)

lin_4_0 <- lm(Survival.Months ~ N.Stage + A.Stage + Tumor.Size + differentiate, data = data_lin)
```

```{r}
# Still prefer lin_4_0
lin_4_1 <- lm(Survival.Months ~ N.Stage + A.Stage + Tumor.Size + Reginol.Node.Positive, data = data_lin)

anova(lin_4_0, lin_4_1)
```

```{r}
# Prefer lin_4_2
lin_4_2 <- lm(Survival.Months ~ N.Stage + A.Stage + Tumor.Size + factor(Estrogen.Status), data = data_lin)

anova(lin_4_0, lin_4_2)
```

```{r}
# Still prefer lin_4_2
lin_4_3 <- lm(Survival.Months ~ N.Stage + A.Stage + Tumor.Size + factor(Estrogen.Status) + factor(Progesterone.Status), data = data_lin)

anova(lin_4_2, lin_4_3)
```
```{r}
par(mfrow = c(2,2))
plot(lin_4_2)
```

### Model Against Status (Logistic Regression)

```{r}
data_log = data |>
  dplyr::select(-Survival.Months, -Regional.Node.Examined, -T.Stage, -Grade) |>
  mutate(Status = as.factor(Status))
```

```{r}
# Basic logistic regression model
full_model <- glm(Status ~ ., data = data_log, family = binomial)
summary(full_model)

stepwise_model <- stepAIC(full_model, direction = "both")
summary(stepwise_model)
```

```{r}
library(car)
vif(full_model)
vif(stepwise_model)
```

```{r}
par(mfrow = c(2, 2))
plot(full_model)
```


```{r}
# Predict probabilities
probabilities <- predict(full_model, type = "response")
```

```{r}
# ROC Curve
library(pROC)
roc_curve <- roc(data_log$Status, probabilities)
auc(roc_curve)
plot(roc_curve)
```




```{r}
#这个地方等之后模型出来了在搞
# Create subsets of data based on race
data_white <- subset(data_log, Race == 0)
data_black <- subset(data_log, Race == 1)

# Fit models on each subset and compare
model_white <- glm(Status ~ ., data = data_white, family = binomial)
model_black <- glm(Status ~ ., data = data_black, family = binomial)
summary(model_white)
summary(model_black)

```